Towards a free/open-source universal-dependency treebank for Kazakh

### 2015-06-20 2015-turklang-kazdep/paper.tex
some minor fixes and notes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60890 72bbbca6-d526-0410-a7d9-f06f51895060
some cleanup of tables
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@60891 72bbbca6-d526-0410-a7d9-f06f51895060

\usepackage{subcaption}

@@

\begin{table}[htbp]
  \centering
\begin{small}
  \begin{tabular}{|l|l|r|r|r|}
    \hline
    \textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
    \hline
    UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
    Phrasebook                     & Phrases from Wikitravel   & 37 & 205 & 5.5 \\
    Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
    Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
    Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
    Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
    Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
    Иран & Wikipedia article (Iran)                                & 111 & 1562 & 14.0 \\
    Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
    \hline
    \multicolumn{2}{c|}{~} & 302 & 3741 & 12.3 \\\cline{3-5}
  \end{tabular}
\end{small}
  \caption{Composition of the corpus. The corpus covers a range of genres and text types from 
    free and public-domain sources.}
\end{table}

@@

\begin{table}[htbp]
  \centering
  \begin{tabular}{|l|l|}
    \hline
    \textbf{Label} & \textbf{Description} \\

    \hline
  \end{tabular}
  \caption{Universal dependency label set}
\end{table}

@@

	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\begin{dependency}[theme = simple, font = \small]
		   \begin{deptext}[column sep=0.3cm]
				Айгүл \& студент \& · \\
				\gmk{np} \& \gmk{n} \& \gmk{cop} \\
				\gloss{Ayğul} \& \gloss{student} \& \gloss{is} \\
			\end{deptext}
			\depedge{2}{1}{\udlabel{subj}}
			\depedge{3}{2}{\udlabel{cop}}
		\end{dependency}
		\caption{xxx }\label{fig:compound}
	\end{subfigure}
	\quad
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\begin{dependency}[theme = simple, font = \small]
		   \begin{deptext}[column sep=0.3cm]
				Айгүл \& студент \& - \\
				\gmk{np} \& \gmk{-} \& \gmk{-} \\
				\gloss{Ayğul} \& \gloss{-} \& \gloss{-}\\
			\end{deptext}
			\depedge{3}{2}{\udlabel{cop}}
		\end{dependency}
		\caption{xxx.}\label{fig:cop2}
	\end{subfigure}
	\quad
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\begin{dependency}[theme = simple, font = \small]
		   \begin{deptext}[column sep=0.3cm]
				Айгүл   \& - \& - \\
				\gmk{n} \& \gmk{-} \& \gmk{-} \& \\
				\gloss{Ayğul} \& \gloss{-} \& \gloss{-} \\
			\end{deptext}
			\depedge{3}{2}{\udlabel{cop}}
		\end{dependency}
		\caption{xxx.}\label{fig:cop3}
	\end{subfigure}

@@

@@

\begin{table}[htbp]
\centering
  \begin{tabular}{|l|l|r|r|}
    \hline
    \textbf{Features}       & \textbf{Algorithm} &\textbf{LAS} & \textbf{UAS} \\
    \hline
     surface                & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
     surface+lemma          & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
     surface+lemma+POS      & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
     surface+lemma+POS+MSD  & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
    \hline
  \end{tabular}
  \caption{Preliminary parsing results from MaltParser using different models. The numbers in 
    brackets denote the upper and lower bounds found during cross validation. Adding 
    structural features to the model substantially improves the performance of the parser, although
    we find no improvement in using full morphosyntactic description (MSD) over using
    simply the first part-of-speech (POS) tag.}
  \label{table:eval}
\end{table}

>>>>

\usepackage{subcaption}
\usepackage{booktabs}
\setlength{\cmidrulewidth}{\heavyrulewidth}

@@

\begin{table}[htbp]
  \centering
\begin{small}
  \begin{tabular}{llrrr}
    \toprule
    \textbf{Document} & \textbf{Description} & \textbf{Sentences} & \textbf{Tokens} & \textbf{Avg. length}\\
    \midrule
    UN Declaration on Human Rights & Legal text on human rights & 25 & 409 & 16.3 \\
    Phrasebook                     & Phrases from Wikitravel   & 37 & 205 & 5.5 \\
    Жиырма Бесінші Сөз             & Philosophical text        & 34 & 525 & 15.4 \\ 
    Қожанасырдың тойға баруы       & Folk tale from Wikisource  & 8 & 140 & 17.5 \\
    Ер Төстік                      & Folk tale from Wikisource  & 23 & 203 & 8.8 \\
    Азамат қайда?                  & Story for language learners & 48 & 434 & 9.0 \\
    Футболдан әлем чемпионаты 2014 & Wikipedia article (2014 World Cup) & 14 & 246 & 17.5 \\
    Иран & Wikipedia article (Iran)                                & 111 & 1562 & 14.0 \\
    Радиан & Wikipedia article (Radian)                            & 2 & 17 & 8.5 \\
    \midrule
    \multicolumn{2}{c}{~} & 302 & 3741 & 12.3 \\\cmidrule{3-5}
  \end{tabular}
\end{small}
  \caption{Composition of the corpus. The corpus covers a range of genres and text types from 
    free and public-domain sources.}
\end{table}

@@

\begin{table}[htbp]
	\centering
	\begin{tabular}{ll}
		\toprule
		\textbf{Label} & \textbf{Description} \\
		\midrule
				\\
 		\bottomrule
	\end{tabular}
	\caption{Universal dependency label set}
\end{table}

@@

	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\begin{dependency}[theme = simple, font = \small]
		   \begin{deptext}[column sep=0.3cm]
				Айгүл \& студент \& · \\
				\gmk{np} \& \gmk{n} \& \gmk{cop} \\
				\gloss{Aygül} \& \gloss{student} \& \gloss{is} \\
			\end{deptext}
			\depedge{2}{1}{\udlabel{subj}}
			\depedge{3}{2}{\udlabel{cop}}
		\end{dependency}
		\caption{xxx }\label{fig:compound}
	\end{subfigure}
	\quad
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\begin{dependency}[theme = simple, font = \small]
		   \begin{deptext}[column sep=0.3cm]
				Айгүл \& студент \& - \\
				\gmk{np} \& \gmk{-} \& \gmk{-} \\
				\gloss{Aygül} \& \gloss{-} \& \gloss{-}\\
			\end{deptext}
			\depedge{3}{2}{\udlabel{cop}}
		\end{dependency}
		\caption{xxx.}\label{fig:cop2}
	\end{subfigure}
	\quad
	\begin{subfigure}[b]{0.3\textwidth}
		\centering
		\begin{dependency}[theme = simple, font = \small]
		   \begin{deptext}[column sep=0.3cm]
				Айгүл   \& - \& - \\
				\gmk{n} \& \gmk{-} \& \gmk{-} \& \\
				\gloss{Aygül} \& \gloss{-} \& \gloss{-} \\
			\end{deptext}
			\depedge{3}{2}{\udlabel{cop}}
		\end{dependency}
		\caption{xxx.}\label{fig:cop3}
	\end{subfigure}

@@

% verbal adjectives → acl
% verbal nouns (gerunds) → ccomp, no xcomp
% verbal adverbs → advlc
% * gna_cnd? → advlc?

@@

\begin{table}[htbp]
	\centering
	\begin{tabular}{llrr}
		\toprule
		\textbf{Features}       & \textbf{Algorithm} &\textbf{LAS} [range] & \textbf{UAS} [range] \\
		\midrule
		surface                & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
		surface+lemma          & nivreeager  & 33.4 [23.4, 42.5] & 50.9 [41.3, 63.2] \\
		surface+lemma+POS      & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
		surface+lemma+POS+MSD  & nivreeager  & 55.2 [43.4, 75.0] & 73.7 [65.8, 92.1] \\
		\bottomrule
	\end{tabular}
	\caption{Preliminary parsing results from MaltParser using different models. The numbers in brackets denote the upper and lower bounds found during cross validation. Adding structural features to the model substantially improves the performance of the parser, although we find no improvement in using full morphosyntactic description (MSD) over using simply the first part-of-speech (POS) tag.}
	\label{table:eval}
\end{table}
